---
output: rmarkdown::github_document
---

```{r echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, 
                      comment = "##", 
                      fig.retina = 2, 
                      fig.path = "README_figs/README-")
```

# Predictive Mean Matched Sequential K-Nearest Neighbor 

## Introduction

The purpose of the this repository is to provide a method for determining the trajectory of any longitudinal outcomes based on obtaining predictions using a extension of a nearest neighbors algorithm described by Dr. Alemi 
(a.k.a. [sequential k-nearest neighbor](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2911789/) (SKNN)). 
We extend the SKNN approach by matching similar patients using the [predictive mean matching](https://amstat.tandfonline.com/doi/abs/10.1080/07350015.1988.10509663). 
We illustrate the use the `PMMSKNN` pacakge below briefly using the `ChickWeight` data.


## Data

The following illustration uses the `ChickWeight` data that exists within `base R`.

## Algorithms Employed 

The main prediction method is using the `R` package [brokenstick](https://github.com/stefvanbuuren/brokenstick), 
along with [predictive mean matching](https://books.google.com/books?hl=en&lr=&id=rM8eSRUYYHYC&oi=fnd&pg=PA442&dq=%22predictive+mean+matching%22++rubin&ots=OM-74mXZoX&sig=H-tIcTl7xqIfbgumXuHBktBTfkQ#v=onepage&q=%22predictive%20mean%20matching%22%20%20rubin&f=false) and 
[gamlss](https://www.gamlss.com/). 
Currently the code is under development to work within the [caret](https://github.com/topepo/caret) and 
[mlr](https://github.com/mlr-org/mlr) packages. 

## Installation/Compilation Tip

- Download the github folder through
``` 
devtools::install_github('ck2136/PMMSKNN')
```

- If not available then `git clone` then `R CMD Install`
```
git clone https://github.com/ck2136/PMMSKNN.git
R CMD Install PMMSKNN
```

- There will be dependencies that should be resolved 
if installation isn't done through the standard R method (in R):
```
devtools::install_github("stefvanbuuren/brokenstick")
devtools::install_deps('.')
devtools::install_local('.')
```

## Example workflow

### Load Libraries and the `ChickWeight` data
```{r package}
library("pacman")
pacman::p_load(PMMSKNN, dplyr, here)
data("ChickWeight") 
## example tug data
```

### Wrangle ChickWeight data
```{r wrangledata}
# Train and Test split for all weight outcome: create 
set.seed(1234)

# load only the TUG dataset
full  <- ChickWeight %>%
  rename(patient_id = 3,
         time = 2) %>% 
  PMMSKNN:::baselinemk(full_0, "patient_id", "time") %>%
  ## filter 10 patients as test with baselinemk()
  mutate(
    patient_id = as.numeric(as.character(patient_id)),
    train_test = ifelse(patient_id %in% c(1, 2, 20, 30, 40), 2, 1),
    Diet = as.numeric(as.character(Diet))) %>% 
  ## add distinct patient id's for the full data
  distinct(patient_id, time, .keep_all = TRUE)
  
## Check the structure of the dataset
# View(full)
attributes(full)$names
attributes(full)$class
```

### PMMSKNN:::preproc() 
creates matched test/train based on Predictive Mean Matching (PMM)
```{r preprocfunc}
test_proc <-
  preproc(
    ## specify full dataset name
    dff = full,
    ## train test split variable
    split_var = "train_test",
    ## training set value as 1
    trainval = 1,
    ## test set value as 2
    testval = 2,
    ## brokenstick knots
    knots_exp = c(0, 4, 8, 16),
    ## specify timepoint to use
    out_time = 16,
    ## specify outcome variable name
    outcome = "weight",
    ## specify time variable name
    time_var = "time",
    ## specify patient id variable name
    pat_id = "patient_id",
    ## Filter observations that will be included
    # filter_exp = "Time > 2",
    ## specify list of covariates for pmm
    varlist = c("Diet"))

use_r("preproc")
test_proc %>% 
  str(max.level=1)
```

- Depending on the `knots_exp` specified and the `out_time` time chosen, there will likely be warnings about parameter estimation within the `brokenstick()` algorithm. Here is where the researcher needs to consider the appropriate values for the variable in terms of clinical relevance and the data at hand.

### LOOCV: `loocv_function()` calculates performance measure 
```{r loocvfunc, results='hide',warning=FALSE, message=FALSE}
res <- loocv_function(
  
  # specify number or vector of numbers from {1,...,total number of patients in training data} 
  nearest_n = c(35:37),
  # enter training and testing post operative and fitted y90 dataset
  train_post = test_proc$train_post,
  ord_data = test_proc$train_o,
  test_post = test_proc$test_post,
  test_o = test_proc$test_o,
  # Specify outcome variable and time variable name
  outcome = "weight",
  time_elapsed = "time",

  # Specify number of cores for parallel processing
  parallel=3,
  
  # Specify use of cubic spline or not
  cs=TRUE,
  
  # specify degrees of freedom use or not
  dfspec=TRUE,
  
  # specify degree of freedom for location, scale and shape (d_f_* where * = {m, s} for location and scale default for shape is 1.
  # specify power transformation of location (ptr_m)
  d_f_m=3, ptr_m=0.5,
  d_f_s=1,
  
  # Specify distribution for location, scale and shape 
  dist_fam = gamlss.dist::NO)
```


### Plots: `plot_cal()` returns a plot of the performance measures from the LOOCV 


```{r plotloocv, warning=FALSE, message=FALSE}
plot_cal(plotobj = res, 
         test_proc = test_proc, 
         obs_dist = "median",
         outcome = "weight",
         filt=FALSE,
         pred_sum="mean",
         #plot_by=seq(10,150,5),
         loocv=TRUE,
         filter_exp = NULL,
         plot_cal_zscore=FALSE,
         wtotplot=FALSE,
         plotvals=FALSE,
         iqrfull=NULL,
         bs=FALSE
         )
```

### Plots: `plot_cal()` also returns plot of the calibration 

```{r plotcal, warning=FALSE, message=FALSE}
plot_cal(plotobj = res, 
         test_proc = test_proc, 
         obs_dist = "median",
         outcome = "weight",
         filt=FALSE,
         pred_sum="mean",
         #plot_by=seq(10,150,5),
         loocv=FALSE,
         filter_exp = NULL,
         plot_cal_zscore=FALSE,
         wtotplot=FALSE,
         plotvals=FALSE,
         iqrfull=NULL,
         bs=FALSE
         )
```


## Authors

* [Chong Hoon Kim](mailto:chong.kim@ucdenver.edu)
* [Dr. Kathryn Colborn](mailto:KATHRYN.COLBORN@UCDENVER.EDU)
* [Dr. Timothy Loar](mailto:TIMOTHY.LOAR@UCDENVER.EDU)
* [Dr. Andrew Kittelson](mailto:andrew.kittelson@ucdenver.edu)
* [Dr. Stef van Buuren](mailto:S.vanBuuren@uu.nl)
